---
title: "Week 2 - Programming in Base R"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

##### 1. Assigned pre measurements and post measurements values to create 2 vectors for blood pressure values

```{r}
pre_bp <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
pre_bp

post_bp <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
post_bp
```

##### 2. Assigned names to the vector elements for both pre and post treatments and for patients 1-20

```{r}
patient <- paste("Patient", 1:20, sep = "_")

names(pre_bp) <- paste("Patient", 1:20, sep = "_")
pre_bp

names(post_bp) <- paste("Patient", 1:20, sep = "_")
post_bp
```

##### 3. Calculated change in blood pressure by subtracting post and pre measurement values and named it "diff_bp"

```{r}
diff_bp <- pre_bp - post_bp
diff_bp
```

##### 4. Calculated average decrease in bp across all patients using the mean function

```{r}
avg_bp_dec <- mean(diff_bp)
avg_bp_dec
```

##### 5. Determined which patients experienced a decrease in bp after treatment (a positive change in bp)

```{r}
patient_bp_dec <- which(diff_bp > 0)
names(patient_bp_dec)
patient_bp_dec
```

##### 6. Subsetted vector to only return those with a positive change by using vex\[index\], returning the actual values that are greater than 0

```{r}
ss_pos_change <- diff_bp[diff_bp > 0]
ss_pos_change
```

##### 7. Calculated the average decrease in blood pressure where the blood pressure decreased (positive change)

```{r}
avg_dec_in_bp <- mean(ss_pos_change)
avg_dec_in_bp
```

## Task 2: Basic Data Frame Practice

##### 1. Created data frame referencing data from task 1

```{r}
patient <- 1:20
pre_bp 
post_bp
diff_bp

r_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
r_df
```

##### 2. Returned only rows of the data where diff_bp is negative by subsetting

```{r}
diff_bp_neg <- subset(patient, diff_bp < 0)
diff_bp_neg
```

##### 3. Added a new column to data frame corresponding to TRUE if post_bp is less than 120 and FALSE if not

```{r}
r_df$post_bp
r_df$post_bp120 <- r_df$post_bp < 120
r_df
```

##### 4. Printed out dataframe, r_df

```{r}
knitr::kable(r_df)
```

## Task 3: List Practice

##### 1. Created a new data frame with new values, looking similar to data frame from task 2 (including new T/F values)

```{r}
patient <- 1:10
pre_plcbo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_plcbo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
diff_plcbo <- pre_plcbo - post_plcbo

plcbo_df <- data.frame(patient, pre_plcbo, post_plcbo, diff_plcbo)
plcbo_df$post_plcbo120 <- plcbo_df$post_plcbo < 120
plcbo_df
```

##### 2. Created and stored a list with 2 elements (1st element = treatment, 2nd element = placebo)

```{r}
treatment <- r_df
placebo <- plcbo_df
list3 <- list(treatment = treatment, placebo = placebo)
list3
```

##### 3. Accessed only the 1st list using 3 different types of syntax: \$, \[\], and \[\[\]\]

```{r}

example1 <- list3$treatment
example1

```

```{r}
bracket <- list3[1]
bracket
```

```{r}
doublebracket <- list3[[1]]
doublebracket
```

##### 4. Used 1 line of code to access the pre_bp (actually referred to as "pre_plcbo") column of the placebo frame (2nd element)

```{r}
list3$placebo$pre_plcbo
```
